Assessing \hydrogen{} and \carbon{} Network Strategies in Europe

### 2023-12-04 report/submissions/journal/paper/paper.tex
start reviewing results section

System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
In the Baseline model, the absence of dedicated carbon and hydrogen transport networks means that the system cannot use all the CC potentials of geographically fixed point sources, but instead needs to deploy more DAC in the locations where carbon is needed for CS and CU.
Naturally, costs for the \carbon{} and \hydrogen{} networks differ between models as they are being deployed.
Finally, we see that the Baseline model deploys more wind and solar than the network models. This is driven by two aspects of the electrolytic hydrogen production: First, the Baseline model deploys more methanation, leading to a larger endogenous hydrogen demand and thus more electricity production. And second, in the network models, the electrolysis is deployed at sites with better renewable resources, thus decreasing the total investments in solar and wind power necessary to run the electrolysis.

@@

We can now closer examine the carbon capture activity when shifting from the \modBase{} to the other three network models.
CU is always deployed in the regions where carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis processes have the lowest cost (or highest value). In the \modBase{}, this means combining the lowest-cost CC source (process emissions) with electrolysis in all regions, even if the renewable resources are poor and electrolysis costs are thus relatively high (e.g., non-coastal regions in Germany). In regions with medium-quality renewables (e.g., France and Scandinavia), CC biomass is also deployed -- and combined with medium-price electrolytic hydrogen -- to run additional CU. But the largest volumes of CU are deployed in the regions with the best renewables and lowest-cost electrolysis (e.g., Iberian Peninsula and British Isles), where all point-source CC potentials are exhausted and additionally necessary carbon is sourced from high-cost DAC.

@@

@@

\subsection*{Hybrid configuration provides flexibilities}

@@

The Hybrid model combines the advantages of both hydrogen and carbon networks. While the topology of the \hydrogen{}-grid roughly corresponds to that of the \hydrogen{} grid model, the \carbon{}-grid plays more of a complementary role. Fig.~\ref{fig:capacity_maps} shows the capacities of the two grids in the hybrid model, with the production and transport capacities given in kt/h. Note that despite similar overall capacities, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.

>>>>

System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
% In the Baseline model, the absence of dedicated carbon and hydrogen transport networks means that the system cannot use all the CC potentials of geographically fixed point sources, but instead needs to deploy more DAC in the locations where carbon is needed for CS and CU.
% Naturally, costs for the \carbon{} and \hydrogen{} networks differ between models as they are being deployed.
% Finally, we see that the Baseline model deploys more wind and solar than the network models. This is driven by two aspects of the electrolytic hydrogen production: First, the Baseline model deploys more methanation, leading to a larger endogenous hydrogen demand and thus more electricity production. And second, in the network models, the electrolysis is deployed at sites with better renewable resources, thus decreasing the total investments in solar and wind power necessary to run the electrolysis.
As we show in the following, these are due to essential though fundamentally different flexibility gains and a shift in carbon utilization and carbon sequestration strategies through the \hydrogen{} and \carbon{} networks.

@@

% This is a bit low-level on regional scale, start with the big picture and dynamics, then zoom in on the regional scale in 3.2 (next section) together with end-use of hydrogen and carbon
% alternative outline:
% 1. baseline model is sub-optimal -> un-used landlocked carbon capture potentials; scarcity in hydrogen and power in Central Europe due to industrial clusters and high population density; costly DAC + additional renewables at sequestration sites
% 2. introduce carbon balance figure -> transport systems unlock point-source CC potentials, in particular biomass and gas for industry, less methanation, less DAC and therefore less DAC specific renewables
% 3. introduce capture share figure -> correlation between average cf and cc share, since gas CHPs and PPs are peak load techs, no cc infrastructure;
% 4. carbon-grid has higher absolute CC and CC shares than hydrogen-grid and less DAC however hydrogen-grid is more cost-effective

We can now closer examine the carbon capture activity when shifting from the \modBase{} to the other three network models.
CU is preferably deployed in the regions where carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis processes have the lowest cost (or highest value). In the \modBase{}, this means combining the lowest-cost CC source (process emissions) with electrolysis in all regions, even if the renewable resources are poor and electrolysis costs are thus relatively high (e.g., non-coastal regions in Germany). In regions with medium-quality renewables (e.g., France and Scandinavia), CC biomass is also deployed -- and combined with medium-price electrolytic hydrogen -- to run additional CU. But the largest volumes of CU are deployed in the regions with the best renewables and lowest-cost electrolysis (e.g., Iberian Peninsula and British Isles), where all point-source CC potentials are exhausted and additionally necessary carbon is sourced from high-cost DAC.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_hydrogen.png}
    \caption{Balance of captured hydrogen across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
    \label{fig:balance_captured_carbon}
\end{figure}

@@

\subsection*{Hybrid configuration provides further flexibilities}

@@

The Hybrid model combines the advantages of both hydrogen and carbon networks. While the topology of the \hydrogen{}-grid roughly corresponds to that of the \hydrogengrid{} model, the \carbon{}-grid plays more of a complementary role. Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two grids in the hybrid model. Note that despite similar overall capacities, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.
